---
title: "IMDB Analysis"
author: "Julian Villar"
date: "30 April 2019"
output: github_document
---

# Are movies getting better?
## Introduction
Do you miss the good old days? Nostalgia really has a way of messing with our perception of whether things are improving. Clearly in technological terms we are better off, but what about culturally? To question whether art is getting better or worse is completely subjective, but what if we could ask everyone what they think on pieces of art throughout time and rate it, that way we could quantify whether there is improvement or not. Thanks to the Internet Movie Database (IMDb) we can.

Using Data acquired that is publicly available from IMDb [here](https://datasets.imdbws.com/), we will report our findings.

## Tools used 
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(ggrepel) 
library(gghighlight)
library(ggridges)
library(tidytext)
library(readr)
library(scales)
library(knitr)
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Ratings
Perhaps before we search through trends throughout time, it is better to establish a baseline. The average movie rating fits this purpose and naturally falls a good entry point to this report.

We introduce our first dataset, `df_ratings`. That is the ratings dataset which we've imported and named `data`. This is our simplest dataset and provides us the Movie ID that’ll act as a Foreign key between our various datasets, the average rating for a movie (averageRating) and the number of votes for each movie (numVotes).

We provide the first few entries to show this dataset clearly.
```{r echo=FALSE, error=FALSE, warning=FALSE, message=FALSE}
read_imdb = function(data_path) {
  path = getwd()
  read_tsv(paste0(path, data_path), na = "\\N", progress=F)
}
df_ratings = read_imdb("\\data.tsv")
df_ratings %>% head()
```
We begin by understanding the spread of average ratings regardless of how many votes each movie has.
```{r error=FALSE, warning=FALSE, message=FALSE}
p1 <- ggplot(df_ratings, aes(x=averageRating)) + 
      geom_histogram(binwidth = 0.4) + 
      labs(x="Count", y="Average Rating", title="Distribution of average rating for all movies")
p1
```
This provides a better understanding than just calculating the mean, strangely we see an average between 6.5-7.5. The implications of this are complex, more than likely we can suggest that people who particularly like certain movies are more likely to vote. However the lack of scores between 2-5 suggest that people really don't use those votes. 

```{r  error=FALSE, warning=FALSE, message=FALSE}
p2 <-     ggplot(df_ratings, aes(x = numVotes, y = averageRating))
p2 <-     p2 +
          geom_bin2d() +
          scale_x_log10(labels = comma) +
          scale_y_continuous(breaks = 1:10) +
          scale_fill_viridis_c(labels = comma,option = "E") +
          labs(x="Number of Votes", y="Average Rating", title="Average IMBd rating and how many people vote that way")
plot(p2)
```


## Expanding our dataset 

We've exhausted this dataset, we'll now extend the data we're looking at to include Movie Titles and their genres/years created. 

```{r error=FALSE, warning=FALSE, message=FALSE}
df_basics = read_imdb("\\data2.tsv")
```
We'll join these together through their unique movie identifier and *improve* our original dataframe.

```{r  error=FALSE, warning=FALSE, message=FALSE}
df_ratings <- df_ratings %>% left_join(df_basics)
head(df_ratings)
```

Now with our new data frame we move to perform the real analysis on our Movie Ratings, particularly with how the Average Rating changes throughout time.

```{r  error=FALSE, warning=FALSE, message=FALSE}
df_ratings_filter = df_ratings %>% filter(titleType == "movie", numVotes >= 10) # Cleaning the Data

p3 <-     ggplot(df_ratings_filter, aes(x = startYear, y = averageRating)) 

p3 <-     p3 + geom_bin2d() +
          geom_smooth(color="black") +
          scale_x_continuous() +
          scale_y_continuous(breaks = 1:10) +
          scale_fill_viridis_c(option = "plasma", labels = comma, trans = 'log10') +
          labs(title="Movie Ratings throughout Time")
plot(p3)
```

As somewhat expected here we see that the average rating for a movie has changed quite dramatically, being just over 5 for movies in the 1930s to almost every movie being around a 7. This probably boils down to recency bias, unless movies really have got drastically better... which I **definitely** won't get into!

Making Horror the subject of this we now need to adapt our code to only include Horror examples. Note that we're not looking for movies such that their genre equals Horror, but rather contains Horror. This is since there exists several movies that have multiple genres! We make us of the fascinating `grepl()` function for this
```{r}
df_ratings_hor = subset(df_ratings, grepl("Horror", genres))
head(df_ratings_hor,15)
```

Now we'll be able to see with the same conditions as before, how the trend looks like for the Horror genre!

```{r  error=FALSE, warning=FALSE, message=FALSE}
df_ratings_hor_filter = df_ratings_hor %>% filter(df_ratings_hor$titleType == "movie", df_ratings_hor$numVotes >= 150) 

p4 <-     ggplot(df_ratings_hor, aes(x = startYear, y = averageRating))

p4 <-     p4 + geom_bin2d() +
          geom_smooth(color="black") +
          scale_x_continuous() +
          scale_y_continuous(breaks = 1:10) +
          scale_fill_viridis_c(option = "plasma", labels = comma, trans = 'log10') + 
          labs(title="Horror movie ratings through time")
plot(p4)

```

Lo and behold, the rating really is going down. We really could say this doesn't look as bad as it really is considering we know there's a potential recency bias in the data. 